---
format: pdf
params:
    my_unitid: 198668
    ipeds_data_year: 20
---


```{r include=FALSE}
ipeds_year <- 2000 + params$ipeds_data_year
file_year <- paste0(as.character(ipeds_year),as.character(params$ipeds_data_year+1))
file_year_dash <- paste0(as.character(ipeds_year),"-",as.character(params$ipeds_data_year+1))
```

---
title: "`r paste0('IPEDS Retention Comparison ', file_year_dash)`" 

---

```{r initialize, include=FALSE, results='asis'}
# title: "`r paste0('IPEDS Retention Comparison ', as.character(2000 + params$ipeds_data_year),'-',as.character(params$ipeds_data_year+1))`" 
library(DBI)
#library(odbc)
library(dplyr)
library(magrittr)
library(tidyr)
#library(stringr)
library(ggplot2)
library(ggdist)

file = paste0("IPEDS", file_year, ".accdb")
connection <- dbConnect(
    odbc::odbc(), 
    .connection_string = paste0(
        "Driver={Microsoft Access Driver (*.mdb, *.accdb)};",
        "Dbq=", file.path(getwd(), file), ";",
        "Uid=Admin;Pwd=;"
      )
)

#mdb_tables_yr <- paste0("Tables", params$ipeds_data_year)
# df_tables <- dbReadTable(connection , "Tables20")

# dbListTables(connection)

source("code/peers.R")

my_peers_ncccs <- get_peer_group(connection, "peer_sets/2-year-NC-public.R") %>%
    select( UNITID, 
            OPEID, 
            INSTNM, 
            INSTSIZE = "(INSTSIZE) Institution size category", 
            C18SZSET = "(C18SZSET) Carnegie Classification 2018: Size and Setting",
            Peer)

my_peers_2year <- get_peer_group(connection, "peer_sets/2-year-public.R") %>%
    select( UNITID, 
            OPEID, 
            INSTNM, 
            INSTSIZE = "(INSTSIZE) Institution size category", 
            C18SZSET = "(C18SZSET) Carnegie Classification 2018: Size and Setting",
            Peer)

mdb_EF_yr <- paste0("EF", ipeds_year)
df_ef <- dbReadTable(connection, paste0(mdb_EF_yr,"D")) %>% 
    select( UNITID,
            FT_Cohort = RRFTCTA,
            FT_Retained = RET_NMF,
            FT_Retention = RET_PCF,
            PT_Cohort = RRPTCTA,
            PT_Retained = RET_NMP,
            PT_Retention = RET_PCP)

df_results_ncccs <- my_peers_ncccs %>%
    left_join( df_ef )

df_results_2year <- my_peers_2year %>%
    left_join( df_ef )

df_results_self <- df_results_ncccs %>%
    filter( UNITID == params$my_unitid )
df_results_ncccs_others <- df_results_ncccs %>%
    filter( UNITID != params$my_unitid )
df_results_2year_others <- df_results_2year %>%
    filter( UNITID != params$my_unitid )


# retention_summary <- df_results_ncccs %>% 
#     mutate(C18SZSET="All NCCCS Schools") %>%
#     group_by(C18SZSET) %>% 
#     summarise(Retention_FT=mean(FT_Retention, na.rm=TRUE),
#               Retention_FT_StdDev=sd(FT_Retention, na.rm=TRUE),
#               Retention_FT_Q25=quantile(FT_Retention, .25, na.rm=TRUE),
#               Retention_FT_Q50=quantile(FT_Retention, .50, na.rm=TRUE),
#               Retention_FT_Q75=quantile(FT_Retention, .75, na.rm=TRUE),
#               Retention_PT=mean(PT_Retention, na.rm=TRUE),
#               Retention_PT_StdDev=sd(PT_Retention, na.rm=TRUE),
#               Retention_PT_Q25=quantile(PT_Retention, .25, na.rm=TRUE),
#               Retention_PT_Q50=quantile(PT_Retention, .50, na.rm=TRUE),
#               Retention_PT_Q75=quantile(PT_Retention, .75, na.rm=TRUE),
#               Schools=n_distinct(UNITID),
#               .groups="keep"
#               ) %>%
#     ungroup() %>%
#     mutate_if(is.numeric, round, 2) %>%
#     mutate(Sort_Val = paste0("ZZZ",C18SZSET),
#            Self = "")
# 
# my_name = df_results_self %$% INSTNM
# 
# retention_summary_grp <- df_results_ncccs %>% 
#     group_by(C18SZSET,Peer) %>% 
#     summarise(Retention_FT=mean(FT_Retention, na.rm=TRUE),
#               Retention_FT_StdDev=sd(FT_Retention, na.rm=TRUE),
#               Retention_FT_Q25=quantile(FT_Retention, .25, na.rm=TRUE),
#               Retention_FT_Q50=quantile(FT_Retention, .50, na.rm=TRUE),
#               Retention_FT_Q75=quantile(FT_Retention, .75, na.rm=TRUE),
#               Retention_PT=mean(PT_Retention, na.rm=TRUE),
#               Retention_PT_StdDev=sd(PT_Retention, na.rm=TRUE),
#               Retention_PT_Q25=quantile(PT_Retention, .25, na.rm=TRUE),
#               Retention_PT_Q50=quantile(PT_Retention, .50, na.rm=TRUE),
#               Retention_PT_Q75=quantile(PT_Retention, .75, na.rm=TRUE),
#               Schools=n_distinct(UNITID),
#               .groups="keep") %>%
#     ungroup() %>%
#     mutate_if(is.numeric, round, 2) %>%
#     mutate(Sort_Val = C18SZSET) %>%
#     mutate_at(vars(contains("_Q")), ~ if_else(Peer, ., NA_real_)) %>%
#     mutate(C18SZSET = case_when(
#                Peer ~ C18SZSET,
#                TRUE ~ my_name),
#            Self = if_else(Peer,"","*")) %>%
# 
#     bind_rows(
#         retention_summary
#     ) %>%
#   
#     arrange(Sort_Val,Self) %>%
#     select(-c(Sort_Val,Peer))

createPlot <- function(df, y, Title) {
    #z <- "FT_Retention"
    y <- enquo(y)
    
    College_Grp <- df_results_self %$% C18SZSET
    College_Ret <- df_results_self %>% summarise(avg=mean(!!y, na.rm=TRUE))

    counts <- df %>% group_by(C18SZSET) %>% tally
    avg_grp <- df %>% group_by(C18SZSET) %>% filter(C18SZSET == College_Grp) %>% summarise(avg=mean(!!y, na.rm=TRUE)) %$% avg
    avg <- df %>% summarise(avg=mean(!!y, na.rm=TRUE)) %$% avg
    
    College_Color <- case_when( # 
        College_Ret >= avg && College_Ret >= avg_grp ~ "Green",
        College_Ret >= avg_grp ~ "Yellow",
        College_Ret >= avg ~ "Orange",
        TRUE ~ "Red"
    )
    
    ggplot(df, aes(x = C18SZSET, y = !!y)) +
        geom_hline(yintercept=avg, color="grey") + # FT_Retention
        ggdist::stat_halfeye(
            adjust = .5,
            width = .6,
            .width = 0,
            justification = -.3,
            point_colour = NA,
            na.rm=TRUE) +
        geom_boxplot(
            width = .25, 
            outlier.shape = NA,
            na.rm=TRUE
          ) +
        scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 15)) +
        geom_point(
            size = 1.3,
            alpha = .3,
            position = position_jitter(
              seed = 1, width = .1
            ),
            na.rm=TRUE
            #,aes(color=Peer)
          ) +
        geom_point(data=df_results_self, aes(x=C18SZSET,y = !!y), color=College_Color, size=5) +
        coord_cartesian(xlim = c(1.2, NA), clip = "off") +
        labs(title = Title, # "Full-time Retention (NCCCS)",
               x = "Institution Type",
               y = "Retention (%)") +
        geom_text(data=counts, aes(x=C18SZSET, y = 3, label=n), size=3)
}

College_Grp <- df_results_self$C18SZSET
College_Name <- df_results_self$INSTNM

dbDisconnect(connection)
```

The following tables and charts are generated from the `r file_year_dash` IPEDS 
Fall Enrollment data. The data is divided into reports on all the NCCCS schools
as well as all 2-year public schools. 

`r College_Name` is identified in the following charts with a large colored dot.
Dot colors are determined according the rules below:

- Green - the college is greater than both the category and overall averages
- Yellow - the college is greater than just the category average
- Orange - the college is greater than just the overall average
- Red - the college is less than both the category and overall averages

The average retention rate for all full-time students at NCCCS and 2-year schools are:

| Category   | Overall Retention (%) | Category Retention (%) |
|:-----------|:---------------------:|:----------------------:|
| All NCCCS  | `r round(mean(df_results_ncccs$FT_Retention, na.rm=TRUE),2)` | `r round(mean(df_results_ncccs[df_results_ncccs$C18SZSET == College_Grp,"FT_Retention"], na.rm=TRUE),2)` |
| All 2-Year | `r round(mean(df_results_2year$FT_Retention, na.rm=TRUE),2)` | `r round(mean(df_results_2year[df_results_2year$C18SZSET == College_Grp,"FT_Retention"], na.rm=TRUE),2)` |
| `r College_Name` | `r df_results_self$FT_Retention`                 | `r df_results_self$FT_Retention` |


The average retention rate for all part-time students at NCCCS and 2-year schools are:

| Category   |  Overall Retention (%) | Category Retention (%) |
|------------|:----------------------:|:----------------------:|
| All NCCCS  | `r round(mean(df_results_ncccs$PT_Retention, na.rm=TRUE),2)` | `r round(mean(df_results_ncccs[df_results_ncccs$C18SZSET == College_Grp,"PT_Retention"], na.rm=TRUE),2)` |
| All 2-Year | `r round(mean(df_results_2year$PT_Retention, na.rm=TRUE),2)` | `r round(mean(df_results_2year[df_results_2year$C18SZSET == College_Grp,"PT_Retention"], na.rm=TRUE),2)` |
| `r College_Name` | `r df_results_self$PT_Retention`                 | `r df_results_self$PT_Retention`                 |


## Full-time Retention Charts

```{r ft_retention_charts, echo=FALSE, warning=FALSE}
df_results_ncccs %>% createPlot(FT_Retention, "Full-time Retention (All NCCCS Schools)")

df_results_2year %>% createPlot(FT_Retention, "Full-time Retention (All 2-year Schools)")
```

# Part-time Retention Charts

```{r pt_retention_charts, echo=FALSE, warning=FALSE}
df_results_ncccs %>% createPlot(PT_Retention, "Part-time Retention (All NCCCS Schools)")

df_results_2year %>% createPlot(PT_Retention, "Part-time Retention (All 2-Year Schools)")
```
